---
title: "Manikin data"
author: "Tina Lasisi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
        - \usepackage[justification=raggedright,labelfont=bf,singlelinecheck=false]{caption}
output:
  pdf_document:
    keep_tex: true
  bookdown::pdf_document2: 
    toc: true
    keep_tex: true
  bookdown::html_document2:
    base_format: workflowr::wflow_html
    toc: true
    toc_float: true
    number_sections: false
    fig_caption: true
editor_options:
  chunk_output_type: console
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(knitr)
library(kableExtra)
library(fs)
library(janitor)
library(fuzzyjoin)
library(tidyquant)
library(ggstatsplot)

# knitting to pdf
# rmarkdown::render(input = "analysis/manikin2021.Rmd", output_format = "pdf_document", output_dir = "output")

F = rprojroot::is_rstudio_project$make_fix_file()

palettedf <- tibble(paletteer::palettes_d_names)

# filter(.data = palettedf, length > 10 & type == "qualitative")

knitr::opts_chunk$set(echo = FALSE, include = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.retina = 2, fig.width = 8, fig.height = 6, out.width = "100%")

```


```{r functions, include=FALSE}

# This chunk contains any functions we may want to use later on in the document.

plot_path = F("output/")

pltsave_func <- function(plot, plot_path, width, height){
  ggsave(
      filename = paste0(deparse(substitute(plot)), ".png"),
      plot = plot, 
      path = plot_path,
      width = width,
      height = height)
  plot(plot)
}

```


```{r df-import, include=FALSE}

# Importing data

Havenith_2020_tidy <- read_csv(F("data/tidy/Havenith_2020_tidy.csv"), 
    col_types = cols(
      wig = col_factor(levels = c("Nude", 
        "LowCurv", "MidCurv", "HighCurv")), 
      radiation = col_factor(levels = c("off", 
            "on")),
      # wind = col_factor(levels = c("0.3", "1", "2.5")), 
      tsk = col_factor(levels = c("34.0", "38.0")),
      wet_dry = col_factor(levels = c("dry", 
            "wet"))))

head(Havenith_2020_tidy)

```


```{r df-summary, include=FALSE}

# means and sd

sum_manikin_df <- Havenith_2020_tidy %>% 
  group_by(wig, radiation, wind, wet_dry, tsk) %>% 
  summarise(across(
    where(is.numeric), 
    list(mean = mean, s = sd), 
    na.rm = TRUE, 
    .names = "{col}_{fn}"
    ),
    n = n()
    )

head(sum_manikin_df)

```


```{r df-solinflux, include=FALSE}

# solar influx

dry_influx_df <- sum_manikin_df %>% 
  select(wig, radiation, wind, wet_dry, tsk, w_m2_mean) %>% 
  filter(wet_dry == "dry") %>% 
  pivot_wider(names_from = radiation, values_from = w_m2_mean) %>% 
  mutate(net_w_m2 = off-on)

corr4c_dry_influx_df <- sum_manikin_df %>% 
  select(wig, radiation, wind, wet_dry, tsk, w_m2_mean, m2k_W_mean) %>% 
  mutate(
    w_m2_4C = 
      case_when(radiation == "off" ~ 28/m2k_W_mean,
                TRUE ~ NaN),
    w_m2_30C = 
      case_when(radiation == "off" ~ 5/m2k_W_mean,
                TRUE ~ NaN)) %>% 
  drop_na() %>% 
  ungroup() %>% 
  select(-radiation)

dry_influx_df2 <- inner_join(dry_influx_df, corr4c_dry_influx_df) %>% 
  mutate(solar_w_m2_4C = w_m2_4C-net_w_m2,
         solar_w_m2_30C = w_m2_30C-net_w_m2,
         sweat_zero_gain = case_when(
           solar_w_m2_30C < 0 ~ abs(solar_w_m2_30C/2430*3600),
           TRUE ~ 0
         )) %>% 
  select(wig, wind, wet_dry, tsk, w_m2_4C, w_m2_30C, solar_w_m2_4C, solar_w_m2_30C, sweat_zero_gain)



wet_influx_df <- sum_manikin_df %>% 
  select(wig, radiation, wind, wet_dry, tsk, w_m2_mean) %>% 
  filter(wet_dry == "wet") %>%
  pivot_wider(names_from = radiation, values_from = w_m2_mean) %>% 
  mutate(net_w_m2 = off-on) %>% 
  drop_na() 

dry_merge <- dry_influx_df2 %>% 
  ungroup() %>% 
  select(-wet_dry, -tsk)

combo_all_df <- full_join(dry_merge, wet_influx_df) %>% 
  mutate(wet_solar_w_m2_30C = w_m2_30C + on,
         wet_w_m2_30C = w_m2_30C + off,
         wet_solar_w_m2_4C = w_m2_4C + on,
         wet_w_m2_4C = w_m2_4C + off,
         diff_dry_wet_30C = wet_solar_w_m2_30C-solar_w_m2_30C,
         sweat_max = diff_dry_wet_30C/2430*3600) %>% 
  select(wig, wind, wet_dry, w_m2_4C, wet_solar_w_m2_4C, wet_w_m2_4C, w_m2_30C, solar_w_m2_30C, wet_solar_w_m2_30C, wet_w_m2_30C, diff_dry_wet_30C, sweat_max)

```

# Heat loss (4C)

The dry measurements were taken at two different temperatures - one with $T_{manikin}= 34^\circ C$ and $T_{ambient}= 8^\circ C$, and another with $T_{manikin}= 38^\circ C$ and $T_{ambient}= 4^\circ C$.

The second set of temperature options was found to be necessary with the straight (low curvature) wig in the $0.3\; m/s$ wind speed setting with the radiation on because the manikin would overheat to the point where no heat loss could be measured. We then made sure to conduct all the $0.3\; m/s$ wind speed experiments with this second option. 

In the preparation of the data, we applied a correction to bring all the measurements to the same temperature, i.e. $T_{ambient}= 4^\circ C$. 

The wet measurements are based on heat exchange rather than dry heat loss. These measurements were taken with the temperature settings $T_{manikin}= 34^\circ C$ and $T_{ambient}= 34^\circ C$ or $T_{manikin}=T_{ambient}$. The measurements have been converted to heat loss for comparative purposes (see Figure \@ref(fig:plt-drywet-scatter4C)).

```{r plt-drywet-scatter4C, fig.cap="Comparison of dry heat loss and wet heat loss (dry + evaporative) for various head coverings at three wind speeds with radiation on and off at 4C."}

dry_influx_df2_4C <- dry_influx_df2 %>% 
  rename(on=solar_w_m2_4C, off=w_m2_4C) %>% 
  select(wig, wind, wet_dry, on, off)

wet_influx_df_4C <- combo_all_df %>% 
  rename(on=wet_solar_w_m2_4C, off = wet_w_m2_4C) %>% 
  select(wig, wind, wet_dry, on, off)

df_4c_influx_wide <- bind_rows(wet_influx_df_4C, dry_influx_df2_4C) %>% 
  mutate(net_w_m2_4C = off-on)

plot <- ggplot(df_4c_influx_wide, aes(off, on)) +
  geom_point(aes(shape = as_factor(wind), fill=wig), size = 5) +
  scale_shape_manual(values = c(21,22,23, 24))+
  theme_bw() +
  geom_abline(slope = 1, intercept = 0)+
  xlim(0, 800) +
  ylim(0, 800) +
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5))+
  labs(x = bquote('Radiation off'~(W/m^2)), y = bquote('Radiation on'~(W/m^2)))+
  guides(fill=guide_legend(override.aes = list(shape=21))) +
  coord_fixed() +
  facet_wrap(vars(wet_dry)) +
  theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin())

print(plot)


```

## Linear model

We created a linear model for heat loss using wet/dry, radiation, wind speed and head covering (wig type) as independent variables. Additionally, we include possible interactions between wet/dry, radiation and wig type.

Below, our results show that the most significant variables are wig and radiation, with marginal significance in radiation and the interaction between radiation and wig type (see also Figure \@ref(fig:plt-lm-heatloss4C)). 

```{r lm-heatloss4C}

df_4c_influx_long <- df_4c_influx_wide %>% 
  group_by(wig, wind, wet_dry) %>% 
  select(-net_w_m2_4C) %>% 
  pivot_longer(on:off, names_to = "radiation", values_to = "heatloss4c")

heatloss4c.lm <- lm(
  formula = 
    heatloss4c ~ wet_dry + radiation + wind + wig + wet_dry*wind + wet_dry*wig + wet_dry*radiation + radiation*wig,
  data = df_4c_influx_long)

summary(heatloss4c.lm)
```

```{r plt-lm-heatloss4C, fig.cap="Dot-and-whisker plot for regression analysis of heatloss with only significant results shown."}
ggstatsplot::ggcoefstats(
  x = heatloss4c.lm,
  sort = "ascending",
  ggtheme = ggplot2::theme_bw(),
  package = "Polychrome",
  palette = "dark",
  only.significant = T
)

```

Viewed as an ANOVA, we confirm that the main significant variables in our model are radiation, wind speed and wig type, while there is a marginal significance (<0.05) for the interaction for radiation * wig type and wet/dry* wig type.

```{r aov-heatloss4C}

summary(aov(heatloss4c.lm))

```


# Solar influx (4C)

We can also look specifically at the effect of the radiation by subtracting the measurements with radiation off from those with radiation on. 

From the plots below, it is apparent that the experiments with a "Nude" manikin scalp show a considerably different pattern than any of the wigs. 

Interestingly, in the dry experiments, the effect of solar radiation appears to cluster more by wig, while the wet experiments show a solar influx that is more patterned by wind speed (see Figure \@ref(fig:plt-drywet-scatter2-4C)).


```{r plt-drywet-scatter2-4C, fig.cap="This plot shows the solar influx as a function of heat loss in the radiation off state. \nThe horizontal line is at zero showing that all values are positive."}


plt_drywet_scatter2_4C <- ggplot(df_4c_influx_wide, aes(off, net_w_m2_4C)) +
  geom_point(aes(shape = as_factor(wind), fill=wig), size = 5) +
  scale_shape_manual(values = c(21,22,23, 24))+
  theme_bw() +
  geom_hline(yintercept =  0) +
  labs(x = bquote('Radiation off'~(W/m^2)), y = bquote('Solar influx'~(W/m^2)))+
  guides(fill=guide_legend(override.aes = list(shape=21))) +
  theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin()) +
  facet_wrap(vars(wet_dry)) 

pltsave_func(plt_drywet_scatter2_4C, plot_path, width = 12, height = 6)


```

Below, we plot the same net heat loss as a function of wind speed. Similarly, we see that, in the dry experiments, there is a very clear effect of wig type and no hair, while the wet experiments show a much more pronounced effect of windspeed (see Figure \@ref(fig:plt-influx4C)).

```{r plt-influx4C, fig.cap="Solar influx as a function of wind speed"}


plt_influx <- ggplot(df_4c_influx_wide, aes(wind, net_w_m2_4C)) +
  geom_point(aes(color=wig, fill=wig), size = 3) +
  ylim(0, 200) +
  geom_path(aes(group=wig, color=wig))+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = bquote('Wind speed'~(m/s)), y=bquote('Solar influx'~(W/m^2))) +
facet_wrap(vars(wet_dry)) +
theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin())

pltsave_func(plt_influx, plot_path, width = 12, height = 6)

```


## Linear model

We test the effect of the predictor variables with a linear model (see Figure \@ref(fig:plt-lm-influx4C)) and find that wig type and wet vs. dry have a significant effect on solar influx, but that wind speed does not.

```{r lm-influx4C}

influx.lm <- lm(formula = net_w_m2_4C ~ wind + wig + wet_dry, 
                data = df_4c_influx_wide)

summary(influx.lm)

```

```{r plt-lm-influx4C, fig.cap="Dot-and-whisker plot of linear regression on solar influx."}

plt_lm_influx <- ggstatsplot::ggcoefstats(
  x = influx.lm,
  sort = "ascending", # sorting the terms of the model based on estimate values
  ggtheme = ggplot2::theme_bw(),
  package = "ggsci",
  palette = "default_aaas",
  messages = F,
  only.significant = T
)

pltsave_func(plt_lm_influx, plot_path, width = 6, height = 10)

```

This is seen more clearly in the ANOVA.

```{r aov-influx4C}

summary(aov(influx.lm))

```

# Heat loss (30C)


```{r plt-drywet-scatter, fig.cap="Comparison of dry heat loss and wet heat loss (dry + evaporative) for various head coverings at three wind speeds with radiation on and off at 30C."}

dry_influx_df2_30C <- dry_influx_df2 %>% 
  rename(on=solar_w_m2_30C, off=w_m2_30C) %>% 
  select(wig, wind, wet_dry, on, off)

wet_influx_df_30C <- combo_all_df %>% 
  rename(on=wet_solar_w_m2_30C, off = wet_w_m2_30C) %>% 
  select(wig, wind, wet_dry, on, off)

df_30C_influx_wide <- bind_rows(wet_influx_df_30C, dry_influx_df2_30C) %>% 
  mutate(net_w_m2_30C = off - on)

plot <- ggplot(df_30C_influx_wide, aes(off, on)) +
  geom_point(aes(shape = as_factor(wind), fill=wig), size = 5) +
  scale_shape_manual(values = c(21,22,23, 24))+
  theme_bw() +
  geom_abline(slope = 1, intercept = 0)+
  xlim(-300, 500) +
  ylim(-300, 500) +
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5))+
  labs(x = bquote('Heat loss with Radiation off'~(W/m^2)), y = bquote('Heat loss with Radiation on'~(W/m^2)))+
  guides(fill=guide_legend(override.aes = list(shape=21))) +
  coord_fixed() +
  facet_wrap(vars(wet_dry)) +
  theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin())

print(plot)


```

## Linear model

We created a linear model for heat loss using wet/dry, radiation, wind speed and head covering (wig type) as independent variables. Additionally, we include possible interactions between wet/dry, radiation and wig type.

Below, our results show that the most significant variables are wig and radiation, with marginal significance in radiation and the interaction between radiation and wig type (see also Figure \@ref(fig:plt-lm-rawheatloss30C)). 

```{r lm-heatloss30C}

df_30C_influx_long <- df_30C_influx_wide %>% 
  group_by(wig, wind, wet_dry) %>% 
  select(-net_w_m2_30C) %>% 
  pivot_longer(on:off, names_to = "radiation", values_to = "heatloss30C")

heatloss30C.lm <- lm(
  formula = 
    heatloss30C ~ wet_dry + radiation + wind + wig + wet_dry*wind + wet_dry*wig + wet_dry*radiation + radiation*wig,
  data = df_30C_influx_long)

summary(heatloss30C.lm)
```

```{r plt-lm-rawheatloss30C, fig.cap="Dot-and-whisker plot for regression analysis of heatloss with only significant results shown."}
ggstatsplot::ggcoefstats(
  x = heatloss30C.lm,
  sort = "ascending",
  ggtheme = ggplot2::theme_bw(),
  package = "Polychrome",
  palette = "dark",
  only.significant = T
)

```

Viewed as an ANOVA, we confirm that the main significant variables in our model are radiation, wind speed and wig type, while there is a marginal significance (<0.05) for the interaction for radiation * wig type and wet/dry* wig type.

```{r aov-rawheatloss30C}

summary(aov(heatloss30C.lm))

```


# Solar influx (30C)

We can also look specifically at the effect of the radiation by subtracting the measurements with radiation off from those with radiation on. 

From the plots below, it is apparent that the experiments with a "Nude" manikin scalp show a considerably different pattern than any of the wigs. 

Interestingly, in the dry experiments, the effect of solar radiation appears to cluster more by wig, while the wet experiments show a solar influx that is more patterned by wind speed (see Figure \@ref(fig:plt-drywet-scatter-zeroline-30C)).


```{r plt-drywet-scatter-zeroline-30C, fig.cap="This plot shows the solar influx as a function of heat loss in the radiation off state. \nThe horizontal line is at zero showing that all values are positive."}


plt_drywet_scatter_zeroline30C <- ggplot(df_30C_influx_wide, aes(off, net_w_m2_30C)) +
  geom_point(aes(shape = as_factor(wind), fill=wig), size = 5) +
  scale_shape_manual(values = c(21,22,23, 24))+
  theme_bw() +
  geom_hline(yintercept =  0) +
  labs(x = bquote('Heatloss Radiation off'~(W/m^2)), y = bquote('Solar influx'~(W/m^2)))+
  guides(fill=guide_legend(override.aes = list(shape=21))) +
  theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin()) +
  facet_wrap(vars(wet_dry)) 

pltsave_func(plt_drywet_scatter_zeroline30C, plot_path, width = 12, height = 6)


```

Below, we plot the same net heat loss as a function of wind speed. Similarly, we see that, in the dry experiments, there is a very clear effect of wig type and no hair, while the wet experiments show a much more pronounced effect of windspeed (see Figure \@ref(fig:plt-influx30C)).

```{r plt-influx30C, fig.cap="Solar influx as a function of wind speed"}


plt_influx30C <- ggplot(df_30C_influx_wide, aes(wind, net_w_m2_30C)) +
  geom_point(aes(color=wig, fill=wig), size = 3) +
  ylim(0, 200) +
  geom_path(aes(group=wig, color=wig))+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = bquote('Wind speed'~(m/s)), y=bquote('Solar influx'~(W/m^2))) +
facet_wrap(vars(wet_dry)) +
theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin())

pltsave_func(plt_influx30C, plot_path, width = 12, height = 6)

```


## Linear model

We test the effect of the predictor variables with a linear model (see Figure \@ref(fig:plt-lm-influx30C)) and find that wig type and wet vs. dry have a significant effect on solar influx, but that wind speed does not.

```{r lm-influx30C}

influx30C.lm <- lm(formula = net_w_m2_30C ~ wind + wig + wet_dry, 
                data = df_30C_influx_wide)

summary(influx30C.lm)

```

```{r plt-lm-influx30C, fig.cap="Dot-and-whisker plot of linear regression on solar influx."}

plt_lm_influx30C <- ggstatsplot::ggcoefstats(
  x = influx30C.lm,
  sort = "ascending", # sorting the terms of the model based on estimate values
  ggtheme = ggplot2::theme_bw(),
  package = "ggsci",
  palette = "default_aaas",
  messages = F,
  only.significant = T
)

pltsave_func(plt_lm_influx30C, plot_path, width = 6, height = 10)

```

This is seen more clearly in the ANOVA.

```{r aov-influx30C}

summary(aov(influx30C.lm))

```

# Effect of radiation and evaporation (30C)

Here are plots for the total heat losses  recalculated for an ambient temperature of $30^\circ C$. 

What becomes apparent now is that there is a substantial heat gain in the dry condition once solar radiation is added (bottom left). In both wet conditions and without the effect of radiation, the absence of hair clearly associates with higher heat loss see Figure \@ref(fig:plt-heat30C)). 


```{r plt-heat30C, fig.width=12, fig.height=12, fig.cap ="Total heat losses with solar exposure \nrecalculated for ambient temperature of 30C"}

plt_heat30_evap_rad <- ggplot(df_30C_influx_long, aes(wind, heatloss30C)) +
# plot <- ggplot(df, aes(wind, heat_loss30)) +
  geom_point(aes(color=wig, fill=wig), size = 3) +
  geom_path(aes(group=wig, color=wig))+
  theme_bw() +
  labs(x = bquote('Wind speed'~(m/s)), y = bquote('Heat loss'~(W/m^2)))+
  facet_wrap(vars(radiation, wet_dry)) +
  ylim(-200, 400)

pltsave_func(plt_heat30_evap_rad, plot_path, width = 12, height = 12)


```

## Linear model

### Separate wet and dry lm

We repeated the linear model for wet and dry conditions separately (see Figure \@ref(fig:plt-lm-drywet30sep)).

We found a striking difference between the effects seen in the dry and wet conditions. The dry condition showed that radiation had the single strongest negative effect on heat loss, while the interaction between solar radiation and high curvature wig showed the highest positive effect on heat loss. All terms were found to be significant. 

Conversely, for the wet conditions, none of the interactions between wig and solar radiation were significant with wind providing the single positive significant effect on heat loss and the wigs reducing heat loss with increased curvature. 

```{r lm-dry30sep}
df <- df_30C_influx_long %>% 
  filter(wet_dry == "dry")

heatloss30C.lm.dry <- lm(
  formula = 
    heatloss30C ~ wind + wig + radiation + radiation*wig, 
  data = df)

summary(heatloss30C.lm.dry) 
```

While the wet condition was most affected by wind speed and saw a negative correlation between heat loss and increasingly curled hair.

```{r lm-wet30sep}
df <- df_30C_influx_long %>% 
  filter(wet_dry == "wet")

heatloss30C.lm.wet <- lm(
  formula = 
    heatloss30C ~ wind + wig + radiation + radiation*wig, 
  data = df)

summary(heatloss30C.lm.wet) 
```


```{r plt-lm-drywet30sep, fig.cap="Dot-and-whisker plot of linear models for heat loss at 30C for (a) dry and (b) wet conditions"}

plt_lm_drywet30sep <-  ggstatsplot::combine_plots(
    ggstatsplot::ggcoefstats(
    x = heatloss30C.lm.dry,
    
    sort = "ascending", # sorting the terms of the model based on estimate values
    ggtheme = ggplot2::theme_bw(),
    package = "ggsci",
    palette = "default_aaas",
    messages = F,
    only.significant = T
  ),
  ggstatsplot::ggcoefstats(
    x = heatloss30C.lm.wet,
    
    sort = "ascending", # sorting the terms of the model based on estimate values
    ggtheme = ggplot2::theme_bw(),
    package = "ggsci",
    palette = "default_aaas",
    messages = F,
    only.significant = T
  ),
  labels = c("a", "b")
  )

pltsave_func(plt_lm_drywet30sep, plot_path, width = 12, height = 6)


```

We present the linear models in an ANOVA table below. The ANOVA for the dry linear model is as follows:

```{r aov-dry30sep}

summary(aov(heatloss30C.lm.dry))

```


For wet, we see: 
```{r aov-wet30sep}

summary(aov(heatloss30C.lm.wet))

```


# Inferred effect of heat loss through sweating at 30C

Here, we plot the sweat rate potential (left) and the sweat rate required to cancel out heat gain at $T_{ambient} = 30^\circ C$.

What emerges is that while heat loss potential is higher without hair as a barrier (i.e. the "nude" condition), the *potential* sweat far exceeds the physiologically possible sweat rate for humans. The plot for zero heat gain shoes that a nude scalp requires the most sweat and this requirement is inversely correlated with hair curvature (see Figure \@ref(fig:plt-sweat)).

```{r df-sweat}
combo1 <- combo_all_df %>% 
  select(wig, wind, wet_dry, sweat=sweat_max)

combo2 <- dry_influx_df2 %>% 
  select(wig, wind, wet_dry, sweat=sweat_zero_gain)

sweat_merge_df <- bind_rows(combo1, combo2) %>% 
  rename(sweat_type=wet_dry) %>% 
  mutate(sweat_type = as_factor(
    case_when(sweat_type=="wet" ~ "max",
              sweat_type=="dry" ~ "zero_gain")))
```


```{r plt-sweat, fig.cap="The quantity of sweat that can be maximally evaporated (left) and that is required for zero heat gain (right) with various head coverings under three wind speeds" ,  fig.width=12}

plt_sweat_max_zero <- ggplot(sweat_merge_df, aes(wind, sweat)) +
  geom_point(aes(color=wig, fill=wig), size = 3) +
  geom_path(aes(group=wig, color=wig))+
  theme_bw() +
  labs(x = bquote('Wind speed'~(m/s)),
       y=bquote('Sweat'~(g/m^2/h))) +
  facet_wrap(vars(sweat_type))

pltsave_func(plt_sweat_max_zero, plot_path, width = 12, height = 8)
```


## Linear model

### Combined wet and dry lm

In a linear model we see that all variables and interactions significantly affect quantity of sweat (see Figure \@ref(fig:plt-lm-sweat)).

```{r lm-sweat}
sweat.lm <- lm(
  formula = 
    sweat ~ sweat_type + wind + wig + sweat_type*wind + sweat_type*wig, 
  data = sweat_merge_df)

summary(sweat.lm) 
```


```{r plt-lm-sweat, fig.cap="Dot-and-whisker plot of linear model for quantity of sweat with wind, wig and sweat type as independent predictors. Only significant values shown"}

ggstatsplot::ggcoefstats(
  x = sweat.lm,
  sort = "ascending", # sorting the terms of the model based on estimate values
  ggtheme = ggplot2::theme_bw(),
  package = "ggsci",
  palette = "default_aaas",
  messages = F,
  only.significant = T
)

```

```{r aov-lm-sweat}

summary(aov(sweat.lm))
```

### Separate max and zero gain

We repeat the tests with maximum sweat potential and zero heat gain separately (see Figure \@ref(fig:plt-lm-sweatsep).

Maximum sweat potential shows a strongly significant effect from wig type and wind speed. 

```{r lm-maxsep}

df <- sweat_merge_df %>% 
  filter(sweat_type == "max")

sweat.lm.max <- lm(
  formula = 
    sweat ~ wind + wig, 
  data = df)

summary(sweat.lm.max) 
```

This is confirmed with an ANOVA.
```{r aov-lm-sweat-max}

summary(aov(sweat.lm.max))
```

Sweat quantity for zero heat gain also shows highly significant effects for wind and wig type.  However, the main difference appears to be in the directionality of the effect of wind.

```{r lm-zerogainsep}
df <- sweat_merge_df %>% 
  filter(sweat_type == "zero_gain") %>% 
  droplevels()

sweat.lm.zero_gain <- lm(
  formula = 
    sweat ~ wind + wig, 
  data = df)

summary(sweat.lm.zero_gain)
```


This is also confirmed with an ANOVA.
```{r aov-lm-sweat-zerogain}

summary(aov(sweat.lm.zero_gain))
```


```{r plt-lm-sweatsep, fig.cap="Dot-and-whisker plots for linear regression of (a) maximum sweat potential and (b) sweat required for zero heat gain"}

plt_lm_sweat_sep <- ggstatsplot::combine_plots(
  ggstatsplot::ggcoefstats(
    x = sweat.lm.max,
    sort = "ascending", # sorting the terms of the model based on estimate values
    ggtheme = ggplot2::theme_bw(),
    package = "ggsci",
    palette = "default_aaas",
    messages = F,
    only.significant = T
  ),
  ggstatsplot::ggcoefstats(
    x = sweat.lm.zero_gain,
    sort = "ascending", # sorting the terms of the model based on estimate values
    ggtheme = ggplot2::theme_bw(),
    package = "ggsci",
    palette = "default_aaas",
    messages = F,
    only.significant = T
  ),
  labels = c("a", "b")
)

pltsave_func(plt_lm_sweat_sep, plot_path, width = 12, height = 8)

```

