---
title: "Thermocouple Summarized Data"
author: "Tina Lasisi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2: default
  bookdown::html_document2:
    base_format: workflowr::wflow_html
    toc: true
    toc_float: true
    number_sections: false
    fig_caption: true
editor_options:
  chunk_output_type: console
---


```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(knitr)
library(tibbletime)
library(lubridate)
library(kableExtra)
library(fs)
library(janitor)
library(fuzzyjoin)
library(timetk)
library(tidyquant)

F = rprojroot::is_rstudio_project$make_fix_file()

# knitting to pdf
# rmarkdown::render(input= "analysis/thermo_sum.Rmd", output_format = "pdf_document", output_dir = "output")

knitr::opts_chunk$set(echo = FALSE, include = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.retina = 2, fig.width = 8, fig.height = 8, out.width = "100%")

```

```{r functions, include=FALSE}

# This chunk contains any functions we may want to use later on in the document.

plot_path = F("output/")

pltsave_func <- function(plot, plot_path, width, height){
  ggsave(
      filename = paste0(deparse(substitute(plot)), ".png"),
      plot = plot, 
      path = plot_path,
      width = width,
      height = height)
  plot(plot)
}

```

# Data Wrangling

First, we import the data directly from the clean thermocouple data (i.e. without outliers). 

```{r read_dfs, include=FALSE}

thermo_df <- F("data/tidy/thermo_data_long_minus_outliers.csv") %>% 
  read_csv()

# converta all variabales with fewer than 5 unique values to factors
col_names <- sapply(thermo_df, function(col) length(unique(col)) < 5)
thermo_df[, col_names] <-lapply(thermo_df[, col_names], factor)
  
head(thermo_df)


```

## Clean 
Here we filter out the trials that used "regular" room conditions with the 0.3m/s setting.

```{r include=FALSE}

merge_df <- as_tbl_time(thermo_df, DateTime) %>%
  pivot_wider(names_from = thermo_id, values_from = temp) %>% 
  filter(!(wind == "0.3" & RoomConditions == "regular"))

```

Then we check for any issues with the data.

```{r}
# check if dataframe contains any N/A
merge_df %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(funs(sum(is.na(.)))) -> extra_NA

head(extra_NA)
```

It appears that thermocouple 1 contains missing data across the trials, so we will remove it. 

```{r include=FALSE}

trim_thermo_df <- merge_df %>% 
  select(-ch1) %>% 
  mutate(seconds_interval = interval(DateTime_start, DateTime),
         seconds_duration = dseconds(int_length(seconds_interval)),
         wig = fct_relevel(wig, "Nude", "LowCurv", "MidCurv", "HighCurv"),
         radiation = fct_relevel(radiation, "off", "on"))

head(trim_thermo_df)

trim_thermo_df %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(funs(sum(is.na(.)))) -> extra_NA

head(extra_NA)

trim_thermo_df_long <- trim_thermo_df %>% 
  pivot_longer(cols = ch2:ch5, names_to = 'thermo_id', values_to = "temp") %>% 
  mutate(thermo_id = as_factor(thermo_id))

head(trim_thermo_df_long)

trim_thermo_df_long %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(funs(sum(is.na(.)))) -> extra_NA

head(extra_NA)

```

# Plots

Below we visualize the cleaned and trimmed data.

## Raw

```{r echo=FALSE}
df <- trim_thermo_df_long
  
wiglist <- unique(levels(df$wig))
windlist <- unique(levels(df$wind))
  
for (Wig in wiglist){
  for (Wind in windlist){
  
  plot <- df %>%
    filter(wind == Wind & wig == Wig) %>% 
    ggplot(
           aes(
             x=seconds_duration, 
             y=temp, 
             group=as.factor(thermo_id), 
             color=thermo_id)) +
    geom_point()+
    geom_path() +
    facet_grid(trial + RoomConditions ~ radiation) +
    ylim(25, 70) +
    theme_bw()+
    ggtitle(paste0(
      "Thermocouple data for ", Wind, " m/s windspeed\n", "for ", Wig)
      ) +
    theme(plot.title = element_text(hjust = 0.5))
  
  print(plot)
  }
}
```



```{r include=FALSE}

mins_thermo_df_long <- trim_thermo_df_long %>% 
  group_by(wig, radiation, wind, trial, thermo_id, RoomConditions) %>% 
  summarize(temp = mean(temp, na.rm = TRUE))

head(mins_thermo_df_long)

mins_thermo_df_long %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(funs(sum(is.na(.)))) -> extra_NA

head(extra_NA)
```


## Thermocouples combined plot

### Radiaton on vs. radiation off

```{r}

scatter_func <- function(df, windspeed){
  df2 <- df %>% 
    dplyr::filter(wind == windspeed)
  
  plot <- ggplot(df2, aes(off, on)) +
    geom_point(aes(shape = thermo_id, fill=wig), size = 5) +
    scale_shape_manual(values = c(21,22,23, 24))+
    theme_bw() +
    ggtitle(paste0(
      "Thermocouple data for ", windspeed, " m/s windspeed\n")
      ) +
    geom_abline(slope = 1, intercept = 0)+
    xlim(25, 60) +
    ylim(25, 60) +
    theme(plot.title = element_text(hjust = 0.5))+
    guides(fill=guide_legend(override.aes = list(shape=21))) +
    coord_fixed()
  
  print(plot)
}

df <- mins_thermo_df_long %>% 
  group_by(wig, radiation, wind) %>% 
  pivot_wider(names_from = radiation, values_from = temp) %>% 
  rowwise() %>% 
  mutate(solar_influx = on-off)

windlist <- unique(levels(df$wind))

for (windspeed in windlist){
  scatter_func(df, windspeed)
}

df_wide <- df %>% 
  pivot_wider(names_from = thermo_id, values_from = c(off, on, solar_influx))

write_csv(df, file = F("data/tidy/summary_thermo_data_long.csv"))
write_csv(df_wide, file = F("data/tidy/summary_thermo_data_wide.csv"))

```

### Solar Influx

```{r}

library(ggbeeswarm)

scatter_func_sol <- function(df){
  df2 <- df 
  
  plot <- ggplot(df2, 
                 aes(
                   color = wig,
                   fill = wig,
                   group = wig,
                   shape = thermo_id,
                   x = wind, 
                   y = solar_influx)) +
    geom_beeswarm(dodge.width = 0.7, size = 3, color="black")+
    scale_shape_manual(values = c(21,22,23, 24))+
    ylim(-10, 20) +
    geom_hline(yintercept = 0) +
    theme_bw() +
    guides(fill=guide_legend(override.aes = list(shape=21)))
  
  print(plot)
}

scatter_func_sol(df)


```

## Thermocouples separate plots

### Radiaton on vs. radiation off

```{r}
scatter_func2 <- function(df, windspeed){
  df2 <- df %>% 
    dplyr::filter(wind == windspeed)
  
  plot <- ggplot(df2, aes(off, on)) +
    xlim(25, 60) +
    ylim(25, 60) +
    geom_point(aes(shape = thermo_id, fill=wig), size = 5) +
    geom_abline(slope = 1, intercept = 0)+
    scale_shape_manual(values = c(21,22,23, 24))+
    coord_fixed() +
    theme_bw() +
    ggtitle(paste0(
      "Thermocouple data for ", windspeed, " m/s windspeed\n")
      ) +
    theme(plot.title = element_text(hjust = 0.5)) +
    facet_wrap(vars(thermo_id)) +
    guides(fill=guide_legend(override.aes = list(shape=21))) 
  
  print(plot)
}

windlist <- unique(levels(df$wind))

for (windspeed in windlist){
  scatter_func2(df, windspeed)
}


```

### Solar Influx

```{r}

scatter_func_sol2 <- function(df){
  df2 <- df 
  
  plot <- ggplot(df2, 
                 aes(
                   color = wig,
                   fill = wig,
                   group = wig,
                   shape = thermo_id,
                   x = wind, 
                   y = solar_influx)) +
    facet_wrap(~thermo_id) +
    geom_beeswarm(dodge.width = 0.7, size = 3, color="black")+
    scale_shape_manual(values = c(21,22,23, 24))+
    ylim(-10, 20) +
    geom_hline(yintercept = 0) +
    theme_bw()+
    guides(fill=guide_legend(override.aes = list(shape=21)))
  
  print(plot)
}

scatter_func_sol2(df)


```


