---
title: "Thermocouple data import and quality control"
author: "Tina Lasisi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: true
    number_sections: yes
editor_options:
  chunk_output_type: console
---


```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(knitr)
library(tibbletime)
library(lubridate)
library(kableExtra)
library(fs)
library(janitor)
library(fuzzyjoin)


F = rprojroot::is_rstudio_project$make_fix_file()

knitr::opts_chunk$set(echo = FALSE, include = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.retina = 2, fig.width = 8, fig.height = 4, out.width = "100%")

```

```{r functions, include=FALSE}

# This chunk contains any functions we may want to use later on in the document.

plot_path = F("output/")

pltsave_func <- function(plot, plot_path, width, height){
  ggsave(
      filename = paste0(deparse(substitute(plot)), ".png"),
      plot = plot, 
      path = plot_path,
      width = width,
      height = height)
  plot(plot)
}

```

# Import data

First, we import the data directly from the raw thermocouple files. 

```{r read_dfs, echo=TRUE}

file_paths <- fs::dir_ls(F("data/raw/ManikinData_Oct2018/ThermoLog/RawThermoFiles/RawThermo_csv/"))


raw_dfs <- file_paths %>% 
  map_dfr(
    read_csv,
    skip = 22,
    na = c("", "NA", "#NA", "#N/A", "Under", "O/C")
  ) %>% 
  mutate(Date = dmy(Date)) %>%
  clean_names() %>% 
  rename_at(.vars = vars(ends_with("_c")),
            .funs = funs(sub("[_]c$", "", .))) %>% 
  select(date:ch5, -type) %>% 
  mutate(DateTime = as_datetime(date(date) + hms(time)))
  
head(raw_dfs)


```

# Import Experiment Log 
Then, we import the log with the times and dates for each experiment (with conditions used).

```{r import_log, echo=TRUE}

# Import the log with each experiment time and conditions
ExperimentLog <- as_tibble(
  read_csv(F("data/tidy/revised_experiment_log.csv"))
) %>% 
  mutate(wig = as_factor(wig),
         radiation = as_factor(radiation),
         trial = as_factor(trial))


head(ExperimentLog)
```

# Clean data 

On November 25th at noon, we had a discussion about the issues with overheating. While we had run a number of trials already, the straight wig with radiation on was the condition where we saw the manikin overheat. According to my notes, we decided that we would *both* lower the room temperature and increase the base temperature of the manikin (to 3 and 37 degrees Celsius, respectively, according to my notes).

Below, I add a number of variables to help visualize including a binary factor of "regular" and "cold" for the room temperature condition (which also would include the higher manikin temperature). The actual temperature of the manikin is included in the dataframe if needed.

```{r clean_log}

# change columns to datetime and create intervals for each trial
DateTimeLog <- ExperimentLog %>% 
  mutate(Trial_Interval = interval(DateTime_start, DateTime_end),
         Trial_Duration = dseconds(int_length(Trial_Interval)),
         RoomConditions = as_factor(
           case_when(mean_amb_temp < 5 ~ "cold",
                     TRUE ~ "regular")),
         wind = as_factor(wind)
  )

# head(DateTimeLog)
```

## Table with log

```{r tbl_log, include=TRUE, results='asis'}

kable(DateTimeLog) %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    fixed_thead = T) %>% 
  scroll_box(width = "600px", height = "500px")

```

Because the duration of the trials was variable, we will have to normalize the duration of the trials for comparison later on. Below is the trial that was the shortest according to these experimental logs.

```{r tbl_log_min, include=TRUE, results='asis'}

DateTimeLog %>% 
  slice(which.min(Trial_Duration)) %>% 
  kable() %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    fixed_thead = T)

```

# Merge data

We can then merge the raw thermocouple data with the log data to create a dataframe where the experimental conditions are associated with the thermocouple readings.

```{r merge_df, echo=TRUE}

ThermoData <- as_tbl_time(raw_dfs, index = DateTime)

# head(ThermoData)

merge_df <- fuzzy_left_join(
  DateTimeLog, ThermoData,
  by = c(
    "DateTime_start" = "DateTime",
    "DateTime_end" = "DateTime"
  ),
  match_fun = list(`<=`, `>=`)
) %>% 
  mutate(seconds_interval = interval(DateTime_start, DateTime),
         seconds_duration = dseconds(int_length(seconds_interval)),
         wig = fct_relevel(wig, "Nude", "LowCurv", "MidCurv", "HighCurv"),
         radiation = fct_relevel(radiation, "off", "on"))


# head(merge_df)

merge_df_long <- merge_df %>% 
  pivot_longer(cols = ch1:ch5, names_to = 'thermo_id', values_to = "temp")

# head(merge_df_long)

```

# Visualize each experimental condition

## All data
The plots show clearly that some of the thermocouples went haywire, registering temperatures in the 900s and above.

Additionally, according to my notes, on November 26th we decided to remove thermocouple 1 from all trials.

```{r plt_thermo_sep, echo=FALSE, fig.height=18, fig.width=12}

df <- merge_df_long %>% 
  mutate(wig = fct_relevel(wig, "Nude", "LowCurv", "MidCurv", "HighCurv"),
         radiation = fct_relevel(radiation, "off", "on"))
  
wiglist <- unique(levels(df$wig))
windlist <- unique(levels(df$wind))
  
for (Wig in wiglist){
  for (Wind in windlist){
  
  plot <- df %>%
    filter(wind == Wind & wig == Wig) %>% 
    ggplot(
           aes(
             x=seconds_duration, 
             y=temp, 
             group=as.factor(thermo_id), 
             color=thermo_id)) +
    geom_point()+
    # geom_line() +
    facet_grid(trial + RoomConditions ~ radiation, scales = "free") +
    theme_bw()+
    ggtitle(paste0(
      "Thermocouple data for ", Wind, " m/s windspeed\n", "for ", Wig)
      ) +
    theme(plot.title = element_text(hjust = 0.5))
  print(plot)
  }
}
  
```


## Extreme temperatures removed
Even with those extremes removed (i.e. temp readings > 100C), it seems that there may be some issues regarding which durations precisely should be used for the thermocouple readings. 

The last 10 minutes of the trial may be bleeding into the next trial, or there may be some other reason for some of the more drastic demarcations within experiments. Additionally, not all the readings are stable for the same amount of time before the end of the experiment.

```{r plt_thermo_sep_filt, echo=FALSE, fig.height=18, fig.width=12}

df <- df %>% 
  filter(temp < 100)
  
wiglist <- unique(levels(df$wig))
windlist <- unique(levels(df$wind))
  
for (Wig in wiglist){
  for (Wind in windlist){
  
  plot <- df %>%
    filter(wind == Wind & wig == Wig) %>% 
    ggplot(
           aes(
             x=seconds_duration, 
             y=temp, 
             group=as.factor(thermo_id), 
             color=thermo_id)) +
    geom_point()+
    # geom_line() +
    facet_grid(trial + RoomConditions ~ radiation, scales = "free") +
    theme_bw()+
    ggtitle(paste0(
      "Thermocouple data for ", Wind, " m/s windspeed\n", "for ", Wig)
      ) +
    theme(plot.title = element_text(hjust = 0.5))
  print(plot)
  }
}

# write_csv(merge_df, file = F("data/tidy/thermo_data_wide.csv"))
# write_csv(merge_df_long, file = F("data/tidy/thermo_data_long.csv"))
# write_csv(df, file = F("data/tidy/thermo_data_long_minus_outliers.csv"))
  
```

## Trimmed data

Based on a visual inspection of the plots, we trim the data to the appropriate time frame and replot.

```{r merge_df2, echo=TRUE}

# Import the log with each experiment time and conditions
ExperimentLog <- as_tibble(
  read_csv(F("data/tidy/trimmed_revised_experiment_log.csv"))
) %>% 
  mutate(wig = 
           as_factor(
             wig, 
             levels = c("Nude", "LowCurv", "MidCurv", "HighCurv")),
         radiation = 
           as_factor(
             radiation,
             levels = c("off", "on")),
         trial = as_factor(trial, ordered = TRUE))


head(ExperimentLog)

# change columns to datetime and create intervals for each trial
DateTimeLog <- ExperimentLog %>% 
  mutate(Trial_Interval = interval(DateTime_start, DateTime_end),
         Trial_Duration = dseconds(int_length(Trial_Interval)),
         RoomConditions = as_factor(
           case_when(mean_amb_temp < 5 ~ "cold",
                     TRUE ~ "regular")),
         wind = as_factor(wind)
  )

head(DateTimeLog)

ThermoData <- as_tbl_time(raw_dfs, index = DateTime)

head(ThermoData)

merge_df <- fuzzy_left_join(
  DateTimeLog, ThermoData,
  by = c(
    "DateTime_start" = "DateTime",
    "DateTime_end" = "DateTime"
  ),
  match_fun = list(`<=`, `>=`)
) %>% 
  mutate(seconds_interval = interval(DateTime_start, DateTime),
         seconds_duration = dseconds(int_length(seconds_interval)))

head(merge_df)

merge_df_long <- merge_df %>% 
  pivot_longer(cols = ch1:ch5, names_to = 'thermo_id', values_to = "temp")

head(merge_df_long)

```



```{r plt_thermo_sep_filt2, echo=FALSE, fig.height=18, fig.width=12}

df <- merge_df_long %>% 
  filter(temp < 100) %>% 
  mutate(wig = fct_relevel(wig, "Nude", "LowCurv", "MidCurv", "HighCurv"),
         radiation = fct_relevel(radiation, "off", "on"))
  
wiglist <- unique(levels(df$wig))
windlist <- unique(levels(df$wind))
  
for (Wig in wiglist){
  for (Wind in windlist){
  
  plot <- df %>%
    filter(wind == Wind & wig == Wig) %>% 
    ggplot(
           aes(
             x=seconds_duration, 
             y=temp, 
             group=as.factor(thermo_id), 
             color=thermo_id)) +
    geom_point()+
    geom_path() +
    facet_grid(trial + RoomConditions ~ radiation) +
    ylim(25, 70) +
    theme_bw()+
    ggtitle(paste0(
      "Thermocouple data for ", Wind, " m/s windspeed\n", "for ", Wig)
      ) +
    theme(plot.title = element_text(hjust = 0.5))
  print(plot)
  }
}

write_csv(merge_df, file = F("data/tidy/thermo_data_wide.csv"))
write_csv(merge_df_long, file = F("data/tidy/thermo_data_long.csv"))
write_csv(df, file = F("data/tidy/thermo_data_long_minus_outliers.csv"))
  
```
